---
title: "jennifer-draft"
author: "Jennifer Tang"
date: "12/10/2019"
output: github_document
---

Libraries
```{r message=FALSE, warning=FALSE}
library(arsenal)
library(dplyr)
library(HH)
```

Importing data
```{r message=FALSE, warning=FALSE}
library(readr)
Lawsuit <- read_csv("data/Lawsuit.csv")
View(Lawsuit)
attach(Lawsuit)
```

Descriptive Statistics
```{r}
dept=as.factor(Lawsuit$Dept)
dept= recode(dept, "1"="Biochemistry/Molecular Biology", "2"="Physiology", 
             "3"="Genetics", "4"="Pediatrics", "5"="Medicine", "6"="Surgery")
gender=as.factor(Lawsuit$Gender)
gender=recode(gender,"1"="Males","0"="Females")
clin=as.factor(Lawsuit$Clin)
clin=recode(clin, "1"="Primarily clinical emphasis", "0"="Primarily research emphasis")
cert= as.factor(Lawsuit$Cert)
cert= recode(cert, "1"="Board Certified", "0"="Not Certified")
rank=as.factor(Lawsuit$Rank)
rank=recode(rank, "1"="Assistant", "2"="Associate", "3"="Full professor")
controls= tableby.control(test=FALSE, total=FALSE,
                          numeric.stats=c("meansd", "medianq1q3"),
                          stats.labels=list(meansd="Mean (SD)", medianq1q3="Median (IQR)"))
tab1= tableby(gender~dept+clin+cert+Prate+Exper+rank+Sal94+Sal95, data=Lawsuit, test=FALSE, 
              total=FALSE, control=controls, digits=2)
summary(tab1, text=TRUE)

```

Distribution of Salary
```{r}
par(mfrow=c(2,2))
hist(Sal94)
hist(Sal95)
trans.sal94=log(Sal94)
trans.sal95= log(Sal95)
hist(trans.sal94)
hist(trans.sal95)

```

Looking at Salary in 1994
``` {r}

###  confounders
mod94=lm(trans.sal94~gender); summary(mod94) 
# gender-dept
summary(lm(trans.sal94~gender+dept))
# gender-clin
summary(lm(trans.sal94~gender+clin))
# gender-cert
summary(lm(trans.sal94~gender+cert))
# gender-Prate
summary(lm(trans.sal94~gender+Prate))
# gender-Exper
summary(lm(trans.sal94~gender+Exper))
# gender- rank
summary(lm(trans.sal94~gender+rank))

#### interactions: dept, experience, rank

#gender-dept
summary(lm(trans.sal94~gender*dept))
# gender-experience 
summary(lm(trans.sal94~gender*Exper))
#gender-rank
summary(lm(trans.sal94~gender*rank))
f1=lm(trans.sal94~dept+clin+cert+Prate+Exper+rank*gender)
summary(f1)

```
Residual plots for model 1994
```{r}
par(mfrow=c(2,2))
plot(f1)
```

Looking at Salary in 1995
```{r}
mod95=lm(trans.sal95~gender);summary(mod95)

# gender-dept
summary(lm(trans.sal95~gender+dept))
# gender-clin
summary(lm(trans.sal95~gender+clin))
# gender-certs
summary(lm(trans.sal95~gender+cert))
# gender-Prate
summary(lm(trans.sal95~gender+Prate))
# gender-Exper
summary(lm(trans.sal95~gender+Exper))
# gender- rank
summary(lm(trans.sal95~gender+rank))
# interaction of dept, exper, rank
summary(lm(trans.sal95~gender*dept)) 
summary(lm(trans.sal95~gender*Exper)) 
summary(lm(trans.sal95~gender*rank)) 

f2=lm(trans.sal95~dept+clin+cert+Prate+Exper+rank*gender)
summary(f2)
```

Residual plots for 1995
```{r}
par(mfrow=c(2,2))
plot(f2)
```

Average Salary of 1994 and 1995
```{r}
mean.salary=(Sal94+Sal95)/2
hist(mean.salary)
log.mean=log(mean.salary)
hist(log.mean)

mean.mod=lm(log.mean~gender); summary(mean.mod)
# there is association between mean salary of 1994 and 1995 and gender

#confounders
summary(lm(log.mean~gender+dept)) # confounder
summary(lm(log.mean~gender+clin)) # confounder
summary(lm(log.mean~gender+cert)) # confounder
summary(lm(log.mean~gender+Prate)) # confounder
summary(lm(log.mean~gender+Exper)) # confounder
summary(lm(log.mean~gender+rank)) # not a confounder

#interactions
summary(lm(log.mean~gender*dept)) # no interaction
summary(lm(log.mean~gender*rank)) # interaction with full professor
summary(lm(log.mean~gender*Exper)) # no interaction

f3= lm(log.mean~dept+clin+cert+Prate+Exper+gender*rank)
summary(f3);
```
 
MULTICOLLINEARITY
suspect multicolinearity of publication rate (based on literature and background research)
```{r}
cor(Lawsuit)
```
high correlation between Prate-clin and prate-dept

#can also check VIF
should drop prate variable

******* final model *********** 
```{r}
f4= lm(log.mean~dept+clin+cert+Exper+gender*rank)
summary(f4)

par(mfrow=c(2,2))
plot(f4)
```

######### outliers and influence points
```{r}
rstandard(f4)
```

```{r}
hat=hatvalues(f4); hat
hat[hat>0.2]
#named numeric(0)
```

```{r}

# remove observation 184 from the mean salary
remove_184=Lawsuit[-c(184),]
log.mean184=(remove_184$Sal94+remove_184$Sal95)/2
dept184=as.factor(remove_184$Dept)
gender184=as.factor(remove_184$Gender)
clin184=as.factor(remove_184$Clin)
cert184= as.factor(remove_184$Cert)
rank184=as.factor(remove_184$Rank)

remove.mod=lm(log.mean184~dept184+clin184+cert184+
                Exper+gender184*rank184, data=remove_184)
summary(remove.mod)
par(mfrow=c(2,2))
plot(remove.mod)
```
There are no outliers for X and Y. Observation 184 is an influential point.
We should keep the influential point in the model since removing data that doesn't necessarily fit the model isn't representative of the general population. Observation 184 is a male not board certified assistant professor who is in the 'Medicine' department. This individual does research primarily and has 2 years of experience with a publication rate of 5.





